---
title: "index.Rmd"
output: html_document
---

<!-- Setup Code. Installs required libaries and reads in dataframe -->
```{r, echo=FALSE, include=FALSE}
library(dplyr)
library(gdata)
library(plotly)
library(knitr)

setwd("~/GitHub/a6-mapping-shootings-arihan-1560795")
shootings.df <- read.csv('data/shootings-2016.csv', stringsAsFactors = FALSE)
```

<!-- Prints summary information from dataset -->
```{r, echo=FALSE, include=FALSE}
#The following chunk of code computes high-level information from the dataset
#Stores the number of shootings we have information about
total.count <- nrow(shootings.df)

#Stores the number of people killed in total
total.dead <- sum(shootings.df$killed)

#Top 5 'affected' cities by number of incidents
top.5.count <- tail(names(sort(table(shootings.df$city))), 5)

#Top 5 cities by number of people killed
city.most.killings <- (aggregate(shootings.df$killed, by=list(Category = shootings.df$city), FUN=sum)) %>% arrange(desc(x))
top.5.dead <- head(city.most.killings, 5)
colnames(top.5.dead) <- c("City", "Total People Killed")
```

`r paste0("**Total number of shootings: **", total.count)`

`r paste0("**Total lives lost in shootings:** ", total.dead)`

`r paste0("**Most affected cities (by number of shootouts): **", (paste(shQuote(top.5.count, type="cmd"), collapse=", ")))`

`r paste0("**Table of most affected cities by sum of people dead in shootouts**")`

`r kable(top.5.dead)`

<!--Summary Table showing sum of people killed over shootouts by state-->
```{r, echo=FALSE, include=FALSE}
sum.state <- (aggregate(shootings.df$killed, by=list(Category = shootings.df$state), FUN=sum))
colnames(sum.state) <- c("State", "Total People Killed")
```
`r kable(sum.state) `

<!-- Description of a particular incident -->
```{r, echo=FALSE, include=FALSE}
#Prints a short description about the worst incident (most number of people killed in one incident).
incident.worst <- shootings.df %>% filter(killed == max(killed))
about.worst <- paste0("On ", incident.worst$date, ", ", incident.worst$killed, " people were killed and ", 
       incident.worst$injured, " others injured in a shootout at a gay nightclub in ", incident.worst$city, 
       ", ", incident.worst$state, ". The gunman, Omar Mateen, 29, of Fort Pierce, Florida allegiance to ISIS
       in a phonecall he made to 911 during the shootout. Omar was interviewed by the FBI in 2013 but was
       not found to be a threat, and his wife claims his actions were a result of being bipolar. 
       After a 3 hour standoff, the police crashed into the building with an armored vehicle
       and stun grenades and killed Omar.")
```
`r about.worst`

`r paste0("Source: [CNN: Orlando shooting: 49 killed, shooter pledged ISIS allegiance](http://www.cnn.com/2016/06/12/us/orlando-nightclub-shooting/)")`

<!--Prints an Interactive Map using information in the provided dataset-->
```{r, echo=FALSE, include=FALSE}
#The following code prints an interactive map of the US with circles representing shootouts in the form of 
#circles. The size of the circles are proptional to the number of people injured and killed in that incident. #Hovering over each circle will overlay additional information about the incident
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)

p <- plot_geo(shootings.df, lat = ~lat, lon = ~lng) %>%
  add_markers(
    text = ~paste0(city, ", ", state, "<br />", 
                   date, "<br />", 
                   paste("Killed:", killed), "<br />", 
                   paste("Injured:", injured)), 
    symbol = I("circle"), 
    hoverinfo = "text", 
    color = "rgb(256, 0, 0)", 
    size = I(as.numeric((shootings.df$killed + shootings.df$injured) * 10))
  ) %>%
  layout(
    title = 'Shootouts across the United States by incident<br />(Hover for details)', geo = g
  )
```
`r p`

<!--Prints an interactive scatterplot-->
```{r, echo=FALSE, include=FALSE}
#The following code prints an interactive scatterplot arranging incidents by when they took place
shootings.df$date <- as.Date(shootings.df$date, format="%B %d, %Y")
shootings.df <- shootings.df %>% mutate(killed + injured)

p1 <- plot_ly(type = 'scatter',
  shootings.df, 
  x = ~date, 
  y = ~(killed + injured),
  color = 'rgb(256,0,0)', 
  size = ~I((killed + injured) * 10),
      text = ~paste0(city, ", ", state, "<br />", 
                     date, "<br />", 
                     paste("Killed:", killed), "<br />", 
                     paste("Injured:", injured)), 
      hoverinfo = "text"
    ) %>% 
  layout(title = "Time vs. Shootouts in the US<br />(Hover for details)", 
         xaxis = list(title = "Date"), 
         yaxis = list(title = "Casualties (People killed + injured)")
         )
```
`r p1`